---
output: github_document
---

```{r rmd-setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "man/figures/")
```

# ClimMobTools

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ClimMobTools)](https://cran.r-project.org/package=ClimMobTools)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/ClimMobTools?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/ClimMobTools)
[![Build Status](https://travis-ci.org/agrobioinfoservices/ClimMobTools.svg?branch=master)](https://travis-ci.org/agrobioinfoservices/ClimMobTools)
[![codecov](https://codecov.io/gh/agrobioinfoservices/ClimMobTools/master.svg)](https://codecov.io/github/agrobioinfoservices/ClimMobTools?branch=master)
[![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
[![Downloads](https://cranlogs.r-pkg.org/badges/ClimMobTools)](https://cran.r-project.org/package=ClimMobTools)
<!-- badges: end -->


Package website: https://agrobioinfoservices.github.io/ClimMobTools/

## Overview

The **ClimMobTools** package provides the toolkit for the [ClimMob](https://climmob.net/climmob3/) platform in R. ClimMob is an open source software for crowdsourcing citizen science in agriculture. Developed by van Etten et al. [(2019)](https://doi.org/10.1017/S0014479716000739), it turns the research paradigm on its head; instead of a few researchers designing complicated trials to compare several technologies in search of the best solutions, it enables many farmers to carry out reasonably simple experiments that taken together can offer even more information. **ClimMobTools** enables project managers to deep explore and analyse their ClimMob data in R.

## Installation

The package may be installed from CRAN via

```{r cran, message=FALSE, eval=FALSE, echo=TRUE}
install.packages("ClimMobTools")
```

The development version can be installed via

```{r install, message=FALSE, eval=FALSE, echo=TRUE}
library("devtools")

devtools::install_github("agrobioinfoservices/ClimMobTools")

```

## Going further

The full functionality of **ClimMobTools** is illustrated in the package vignette. The vignette can be found on the [package website](https://agrobioinfoservices.github.io/ClimMobTools/) or from within `R` once the package has been installed, e.g. via 

```{r vignette, message=FALSE, eval=FALSE, echo=TRUE}

vignette("Overview", package = "ClimMobTools")

```


## Contribution

You are welcome to contribute to this project. Please read our [contribution guide lines](CONTRIBUTING.md). 

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.
