---
title: "Introduction to ClimMobTools"
package: ClimMobTools
author:
- name: Kauê de Sousa
  affiliation: Department of Agricultural Sciences, Høgskolen i Innlandet, Norway; and Bioversity International
- name: Jacob van Etten
  affiliation: Bioversity International, Rome, Italy
- name: Sam Dumble
  affiliation: Statistics for Sustainable Development, Reading, UK
- name: Brandon Madriz
  affiliation: Bioversity International, Turrialba, Costa Rica
- name: Carlos F. Quiros
  affiliation: Bioversity International, Turrialba, Costa Rica
output: html_document
vignette: >
  %\VignetteIndexEntry{Introduction to ClimMobTools}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: ["ClimMobTools.bib"]
csl: citation_style.csl
---

```{r setup, include=FALSE}
TRAVIS <- !identical(tolower(Sys.getenv("TRAVIS")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = TRAVIS
)
```

## Overview

The **ClimMobTools** package provides the toolkit employed in crowdsourcing citizen science projects under the *tricot* approach. Tricot, stands for "triadic comparison of technologies", an approach developed by van Etten et al. (2016)[@VanEtten2016] for the rapid assessment of on-farm evaluation trails in small-scale agriculture.

## Installation

The development version can be installed via


    install.packages("devtools")
    devtools::install_github("agrobioinfoservices/ClimMobTools", upgrade = "never")
    
    
# Usage

The `breadwheat` is a dataframe from crowdsourcing citizen-science trials of bread wheat (*Triticum aestivum*) varieties in India. This is a sample data available at the [ClimMob](https://climmob.net/blog/) that can be fetched using using the function `getDataCM` from **ClimMobTools** and an API key from the ClimMob portal.

```{r fetch, message=FALSE, eval=FALSE, echo=TRUE}
library("ClimMobTools")
# library("devtools")
# devtools::install_github("agrobioinfoservices/gosset", upgrade = "never")
library("gosset")
library("PlackettLuce")
library("tidyverse")
library("magrittr")

# the API key
key <- "d39a3c66-5822-4930-a9d4-50e7da041e77"

data <- ClimMobTools::getDataCM(key = key, 
                                project = "breadwheat")


```


## Preparing the data

The function `getDataCM` returns a dataframe in the long format to allows the organisation of projects with different structure. We can put this into wide format where each row is a farmer observation using **tidyverse**. 


```{r tidying, eval=FALSE, echo=TRUE}
# reshape the data into the wide format
# we also remove some variables that refer to the ODK survey employed for the data collection
data %<>% 
  filter(!str_detect(variable, "survey")) %>% 
  group_by(id) %>%
  distinct(id, variable, value) %>%
  spread(variable, value)
```

The [PlackettLuce](https://hturner.github.io/PlackettLuce/) model is one approach to analyse the *tricot* data[@Turner2018]. To do so, we need to convert the farmers' ranking into a object of class 'rankings' or 'grouped_rankings'. A object of class 'rankings' is a matrix of dense rankings that can be used to fit a PlackettLuce model using `PlackettLuce` from **PlackettLuce**. A object of class 'grouped_rankings' associates a group index with an object of class 'rankings', then it allows the rankings to be linked to explanatory variables and fit a PlackettLuce model using `pltree` from **PlackettLuce**. 

We can convert the farmers' rankings in into a rankings object using the function `to_rankings` from package **gosset**. The argument `grouped.rankings` return a object of class 'grouped_rankings'.

```{r plrankings, message = FALSE, eval=FALSE, echo=TRUE}

G <- to_rankings(data, 
                 items = c("item_A","item_B","item_C"), 
                 input = c("overallperf_pos","overallperf_neg"),
                 type = "tricot", 
                 grouped.rankings = TRUE)


```


## Add environmental variables

Environmental variables can easily be included into the data set using the longitude and latitude information from the **tricot** and the planting dates for each plot. In this exercise we include the heat stress indices using the function `temperature` from **ClimMobTools**.

```{r temperature, message=FALSE, eval=FALSE, echo=TRUE}

# first we convert the lon lat into numeric
# and the planting dates into Date
data %<>%
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat),
         plantingdate = as.Date(plantingdate, 
                                format = "%Y-%m-%d"))

# then we get the temperature indices
# get some variables to include in the model
# from the planting date to the next 120 days
temp <- temperature(data[c("lon","lat")], 
                    day.one = data["plantingdate"], 
                    span = 120)


modeldata <- cbind(G, temp)


```

# Fit a PlackettLuce model

Now we fit the model using the function `pltree` from **PlackettLuce** with temperature indices as explanatory variables.

```{r plmodel, message=FALSE, eval=FALSE, echo=TRUE}
tree <- pltree(G ~ ., data = modeldata, npseudo = 5)

```

The PlackettLuce model shows that the bread wheat variaties had a different performance under a threshold of 15.5 degrees Celsius for the diurnal temperature range (DTR).  


# References




