---
title: "ClimMobTools: API Client for the 'ClimMob' platform in R"
package: ClimMobTools
author:
- name: Kauê de Sousa
  affiliation: Department of Agricultural Sciences, Inland Norway University of Applied Sciences, Hamar, Norway <br> Digital Inclusion, Bioversity International, Montpelier, France
- name: Jacob van Etten
  affiliation: Digital Inclusion, Bioversity International, Montpelier, France
- name: Brandon Madriz
  affiliation: MrBot Software Solutions, Cartago, Costa Rica
output: html_document
vignette: >
  %\VignetteIndexEntry{ClimMobTools}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: ["ClimMobTools.bib"]
csl: citation_style.csl
---

```{r setup, include=FALSE}
TRAVIS <- !identical(tolower(Sys.getenv("TRAVIS")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = TRAVIS
)
```

## Overview

**ClimMobTools** the API Client for the 'ClimMob' platform in R. [ClimMob](https://climmob.net/) is an open source software for crowdsourcing citizen science in agriculture under the tricot approach[@vanEtten2019tricot]. The approach aims the rapid assessment of technologies in the target environment. Tricot turns the research paradigm on its head; instead of a few researchers designing complicated trials to compare several technologies in search of the best solutions, it enables many participants to carry out reasonably simple experiments that taken together can offer even more information.

# Usage

The `breadwheat` is a dataframe from crowdsourcing citizen science trials of bread wheat (*Triticum aestivum* L.) varieties in India. This is a sample data available at the [ClimMob](https://climmob.net/) that can be fetched using the function `getDataCM()` from **ClimMobTools** and an API key from the ClimMob user's account.

```{r fetch, message=FALSE, eval=FALSE, echo=TRUE}
library("ClimMobTools")

# the API key
key <- "d39a3c66-5822-4930-a9d4-50e7da041e77"

dat <- getDataCM(key = key,
                 project = "breadwheat",
                 pivot.wider = TRUE)


names(dat) <- gsub("firstassessment_|package_|lastassessment_|registration_", "",
                   names(dat))

```

# Tricot data with environmental covariates

We can add environmental covariates from package [climatrends](https://CRAN.R-project.org/package=climatrends). Here we use function `temperature()` to compute the temperature indices for the first 80 days after planting. 

```{r temperature, message=FALSE, eval=FALSE, echo=TRUE}
library("climatrends")
library("nasapower")

dat$plantingdate <- as.Date(dat$plantingdate, format = "%Y-%m-%d")
dat$lon <- as.numeric(dat$farm_geo_longitude)
dat$lat <- as.numeric(dat$farm_geo_latitude)

temp <- temperature(dat[, c("lon","lat")], 
                    day.one = dat[, "plantingdate"],
                    span = 80)

temp
```


# Tricot data into rankings

The Plackett-Luce model is one approach to analyse the ClimMob data[@Turner2020]. We build the farmers' rankings as an object of class 'grouped_rankings'. This allows the rankings to be linked to the environmental covariates computed previously and fit the model using `pltree()` from the package [PlackettLuce](https://CRAN.R-project.org/package=PlackettLuce). 

We build the rankings using the function `rankTricot()`. 

```{r plrankings, message=FALSE, eval=FALSE, echo=TRUE}
library("PlackettLuce")

R <- rankTricot(dat, 
                items = c("item_A","item_B","item_C"), 
                input = c("overallperf_pos","overallperf_neg"),
                group = TRUE)

pld <- cbind(R, temp)

pl <- pltree(R ~ maxNT + maxDT, 
             data = pld)

summary(pl)

plot(pl)

```


# References


