---
title: "ClimMobTools: API Client for the 'ClimMob' platform in R"
package: ClimMobTools
author:
- name: Kauê de Sousa
  affiliation: Digital Inclusion, Bioversity International, Montpellier, France <br> Department of Agricultural Sciences, University of Inland Norway, Hamar, Norway
- name: Jacob van Etten
  affiliation: Digital Inclusion, Bioversity International, Montpelier, France
- name: Brandon Madriz
  affiliation: MrBot Software Solutions, Cartago, Costa Rica
output: html_document
vignette: >
  %\VignetteIndexEntry{ClimMobTools}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{PlackettLuce}
bibliography: ["ClimMobTools.bib"]
csl: citation_style.csl
---

```{r setup, include=FALSE}
library("ClimMobTools")
library("PlackettLuce")
```

## Overview

**ClimMobTools** is the API client for the 'ClimMob' platform in R. [ClimMob](https://climmob.net/) is open-source software for decentralized large-N trials using the 'tricot' approach[@vanEtten2019tricot]. This approach facilitates the rapid assessment of technologies in target environments. Tricot turns the research paradigm on its head: instead of a few researchers designing complicated trials to compare several technologies in search of the best solutions, it enables many participants to carry out reasonably simple experiments that, taken together, can offer even more information.

## Usage

The `breadwheat` dataset is a dataframe from crowdsourced citizen science trials of bread wheat (*Triticum aestivum* L.) varieties in India. This sample data is available on the [ClimMob](https://climmob.net/) platform and can be fetched using the `getDataCM()` function from **ClimMobTools**, along with an API key from the ClimMob user's account.

```{r fetch, message=FALSE, eval=TRUE, echo=TRUE}
library("ClimMobTools")
library("PlackettLuce")

# the API key
key <- "d39a3c66-5822-4930-a9d4-50e7da041e77"

dat <- getDataCM(key = key,
                 project = "breadwheat",
                 userowner = "gosset",
                 pivot.wider = TRUE)


names(dat) <- gsub("firstassessment_|package_|lastassessment_|registration_", "",
                   names(dat))

```

# Tricot data into rankings

The Plackett-Luce model is one approach to analyze the ClimMob data[@Turner2020]. We build the farmers' rankings as an object of class `rankings` from the package [PlackettLuce](https://CRAN.R-project.org/package=PlackettLuce). We build the rankings using the function `rankTricot()`.

```{r plrankings, message=FALSE, eval=TRUE, echo=TRUE}
R = rankTricot(dat, 
                items = c("item_A","item_B","item_C"), 
                input = c("overallperf_pos","overallperf_neg"))

mod = PlackettLuce(R)

summary(mod)
```

For more analytical insights, please visit the documentation of the [gosset](https://cran.r-project.org/package=gosset) package.

# References


