---
title: "Introduction to ClimMobTools"
package: ClimMobTools
author:
- name: Kauê de Sousa
  affiliation: Department of Agricultural Sciences, Høgskolen i Innlandet, Hamar, Norway; and Bioversity International
- name: Jacob van Etten
  affiliation: Bioversity International, Rome, Italy
- name: Sam Dumble
  affiliation: Statistics for Sustainable Development, Reading, UK
- name: Brandon Madriz
  affiliation: Bioversity International, Turrialba, Costa Rica
- name: Carlos F. Quiros
  affiliation: Bioversity International, Turrialba, Costa Rica
output: html_document
vignette: >
  %\VignetteIndexEntry{Introduction to ClimMobTools}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: ["ClimMobTools.bib"]
csl: citation_style.csl
---

```{r setup, include=FALSE}
TRAVIS <- !identical(tolower(Sys.getenv("TRAVIS")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = TRAVIS
)
```

## Overview

The **ClimMobTools** package provides the toolkit employed in crowdsourcing citizen science projects under the *tricot* approach. Tricot, stands for "triadic comparison of technologies", an approach developed by van Etten et al. (2016)[@VanEtten2016] for the rapid assessment of on-farm evaluation trails in small-scale agriculture.

## Installation

The development version can be installed via


    library("devtools")
    devtools::install_github("agrobioinfoservices/ClimMobTools", upgrade = "never")
    
    
# Usage

The `breadwheat` is a dataframe from crowdsourcing citizen science trials of bread wheat (*Triticum aestivum*) varieties in India. This is a sample data available at the [ClimMob](https://climmob.net/blog/) that can be fetched using the function `getDataCM` from **ClimMobTools** and an API key from the ClimMob portal.

```{r fetch, message=FALSE, eval=FALSE, echo=TRUE}
library("ClimMobTools")
library("tidyverse")
library("magrittr")

# the API key
key <- "d39a3c66-5822-4930-a9d4-50e7da041e77"

data <- ClimMobTools::getDataCM(key = key, 
                                project = "breadwheat",
                                tidynames = TRUE)

# reshape the data into the wide format
data %<>% 
  filter(!str_detect(variable, "survey")) %>% 
  group_by(id) %>%
  distinct(id, variable, value) %>%
  spread(variable, value) %>% 
  ungroup()


```

## Environmental variables

Understanding how different crop varieties respond to environmental variation is a valuable approach to provide recommendations based on genotype x environment (GxE) interactions for climate adaptation. **ClimMobTools** provides the toolkit to compute environmental variables that serves as covariates for the ranking analysis with explanatory variables using  [Plackett-Luce](https://CRAN.R-project.org/package=PlackettLuce)[@Turner2018] or [Bradley-Terry](https://CRAN.R-project.org/package=psychotree)[@Wickelmaier2018] model. 

The climatic variables available in **ClimMobTools** were previously used for the environmental analysis of wheat trials[@Kehel2016] and can explain the changes in climatic patterns due to climate change[@Aguilar2005]. In a more advanced analysis, these variables can be applied with seasonal forecasts and risk analysis to provide varietal recommendations for climate adaptation[@VanEtten2019].

The environmental data can be calculated using input from time series databases such as the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS)[@Funk2015] for rainfall, and MODIS (MYD11A2)[@Wan2015] for land surface temperature. These databases has the advantage to be free and publicly available sources of global coverage and high resolution (0.05°), and have been used in previous studies with ClimMob data[@VanEtten2019]. However, it requires much computer capacity to extract this information and produce the matrix for the environmental indices. MODIS data also requires some work to reduce noise and fill gaps.

The other alternative relies on using [NASA POWER](https://power.larc.nasa.gov/). By providing the planting dates and the geographic information (lonlat), **ClimMobTools** makes an internal call to [nasapower](https://CRAN.R-project.org/package=nasapower)[@Sparks2018] to fetch the required time series data and calculate the environmental indices. It do not requires much computer capacity but requires a persistent internet connection. NASAPOWER, however has a low grid resolution (0.5°) compared to MODIS and CHIRPS and could not provide the expected results for GxE interactions in a narrow geographical range.

### Temperature

**ClimMobTools** functions to compute environmental indices has the basics input data; a numeric vector of geographic information (lonlat) or an array with day and night temperature (from sources like MODIS) and the a vector of class `Date` for the first day that will be taken into account for the environmental indices. Generally the `Date` vector is the planting date from when the experiment were established, but can also refer to other phenological event during the plant development, such as day of 50% flowering or the first day of a growing season. 

The duration from where the environmental indices will be computed is defined by the argument `span` which can be a single integer that takes into account a single time span for the entire data set or a vector indicating the time span for each observation. 

Here we calculate the temperature indices for a time span of 120 days after the planting date using the function `temperature` from **ClimMobTools**.

```{r temperature, message=FALSE, eval=FALSE, echo=TRUE}

# first we convert the lon lat into numeric
# and the planting dates into Date
data %<>%
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat),
         plantingdate = as.Date(plantingdate, 
                                format = "%Y-%m-%d"))

# then we get the temperature indices
temp <- temperature(data[c("lon","lat")], 
                    day.one = data["plantingdate"], 
                    span = 120)


```

### Rainfall 

Precipitation indices can also be calculated in the same way as the temperature indices. The can be either the geographic coordinates or a matrix with CHIRPS data.

Here we use the NASA POWER data with a time span of 120 days after planting date using the function `rainfall` from **ClimMobTools**.  

```{r rain, message=FALSE, eval=FALSE, echo=TRUE}
rain <- rainfall(data[c("lon","lat")], 
                 day.one = data["plantingdate"],
                 span = 120)

```

Function `rainfall` has also a feature that enables the indices to be calculated from some days before the planting date. This is important for some studies where a residual precipitation must be taken into account. To do so, we use the argument `days.before`:


```{r rain2, message=FALSE, eval=FALSE, echo=TRUE}
rain <- rainfall(data[c("lon","lat")], 
                 day.one = data["plantingdate"], 
                 span = 120,
                 days.before = 15)


```

# Model with environmental variables

The Plackett-Luce model is one approach to analyse the ClimMob data[@Turner2018]. To do so, we need to convert the farmers' ranking into an object of class 'grouped_rankings'. An object of class 'grouped_rankings'  allows the rankings to be linked to explanatory variables and fit the model using `pltree` from **PlackettLuce**. 

We can convert the farmers' rankings in into a grouped_rankings object using the function `to_rankings` from package **gosset**. 

```{r plrankings, message=FALSE, eval=FALSE, echo=TRUE}
G <- build_rankings(data, 
                    items = c("item_A","item_B","item_C"), 
                    input = c("overallperf_pos","overallperf_neg"),
                    grouped.rankings = TRUE)


```


Finally, we fit the model using the function `pltree` from **PlackettLuce** with temperature and rainfall indices as explanatory variables.

```{r plmodel, message=FALSE, eval=FALSE, echo=TRUE}
library("PlackettLuce")

modeldata <- cbind(G, temp, rain)

tree <- pltree(G ~ ., data = modeldata, npseudo = 5)

```

The PlackettLuce model shows that the bread wheat varieties had a different performance under a threshold of 15.5 degrees Celsius for the diurnal temperature range (DTR).  

# References




