---
title: "ClimMobTools: Toolkit for the 'ClimMob' platform in R"
package: ClimMobTools
author:
- name: Kauê de Sousa
  affiliation: Department of Agricultural Sciences, Inland Norway University of Applied Sciences, Hamar, Norway; and Bioversity International, Rome, Italy
- name: Jacob van Etten
  affiliation: Bioversity International, Rome, Italy
- name: Brandon Madriz
  affiliation: Bioversity International, Turrialba, Costa Rica
- name: Carlos F. Quiros
  affiliation: Bioversity International, Turrialba, Costa Rica
output: html_document
vignette: >
  %\VignetteIndexEntry{ClimMobTools}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: ["ClimMobTools.bib"]
csl: citation_style.csl
---

```{r setup, include=FALSE}
TRAVIS <- !identical(tolower(Sys.getenv("TRAVIS")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = TRAVIS
)
```

## Overview

**ClimMobTools** the API for the 'ClimMob' platform in R. [ClimMob](https://climmob.net/) is an open source software for crowdsourcing citizen science in agriculture under the tricot approach[@vanEtten2019tricot]. The approach aims the rapid assessment of on-farm evaluation trails in small-scale agriculture. Tricot turns the research paradigm on its head; instead of a few researchers designing complicated trials to compare several technologies in search of the best solutions, it enables many farmers to carry out reasonably simple experiments that taken together can offer even more information.

## Installation

The package may be installed from CRAN via

```{r cran, message=FALSE, eval=FALSE, echo=TRUE}
install.packages("ClimMobTools")
```

The development version can be installed via

```{r install, message=FALSE, eval=FALSE, echo=TRUE}
library("devtools")

devtools::install_github("agrobioinfoservices/ClimMobTools")

```


# Usage

The `breadwheat` is a dataframe from crowdsourcing citizen science trials of bread wheat (*Triticum aestivum* L.) varieties in India. This is a sample data available at the [ClimMob](https://climmob.net/) that can be fetched using the function `getDataCM` from **ClimMobTools** and an API key from the ClimMob user's account.

```{r fetch, message=FALSE, eval=FALSE, echo=TRUE}
library("ClimMobTools")
library("tidyverse")
library("magrittr")

# the API key
key <- "d39a3c66-5822-4930-a9d4-50e7da041e77"

data <- getDataCM(key = key, 
                  project = "breadwheat",
                  tidynames = TRUE)

# reshape the data into the wide format
data %<>% 
  filter(!str_detect(variable, "survey")) %>% 
  group_by(id) %>%
  distinct(id, variable, value) %>%
  spread(variable, value) %>% 
  ungroup()


```

# Tricot data into rankings

The Plackett-Luce model is one approach to analyse the ClimMob data[@Turner2018]. To do so, we build the farmers' rankings as an object of class 'grouped_rankings'. This allows the rankings to be linked to explanatory variables and fit the model using `pltree` from **PlackettLuce**. 

We build the rankings using the function `rank_tricot` from the R package **gosset**. 

```{r plrankings, message=FALSE, eval=FALSE, echo=TRUE}
library("PlackettLuce")
library("gosset")

G <- rank_tricot(data, 
                 items = c("item_A","item_B","item_C"), 
                 input = c("overallperf_pos","overallperf_neg"),
                 group = TRUE)

plt <- pltree(G ~ .)


```


# References




